---
title: "Lab 4 Assignment"
author: "Dhananjay"
date: "2023-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Q1. Compare Bar and Pie Plots
Consider the dataset below.  Please construct two plots to visualize this data: a sorted bar plot and a pie chart.  Make sure both plots have easy to read labels and grid lines for reference.
http://eecs.ucf.edu/~wiegand/idc6700/datasets/doctorwho.csv

```{r}

# Loading the required libraries
library(tidyverse)
library(ggplot2)

# Loading the doctorwho dataset
doctorWhodf <- read.csv("doctorwho.csv", header = TRUE, sep = ",")

# Computing the frequency of each doctor
doctorFrequency <- table(doctorWhodf$FavoriteDoctor)

# Sorting the frequency in descending order
sortedDoctorFrequency <- sort(doctorFrequency, decreasing = TRUE)

# Creating a vector of colors for each doctor
doctorColors <- c("blue", "red", "black", "purple")

# creating a sorted bar plot with different colors and grid lines
doctor_bp <- barplot(sortedDoctorFrequency, 
              main = "Patient's Visiting Frequency per Doctors", 
              xlab = "Doctors",
              ylab = "Frequency", 
              col = doctorColors, 
              border = "white", 
              ylim = c(0, 25),
              las = 1, 
              cex.names = 1, 
              horiz = FALSE, 
              yaxs = "i")

# Adding a horizontal grid line at the x-axis
abline(h = 0, v = NULL, col = "black", lty = "solid")

# Adding horizontal grid lines at each y-interval
abline(h = seq(0, 20, by = 5), v = NULL, lty = "dotted", col = "gray")

# Adding frequency values on top of each bar
text(doctor_bp, sortedDoctorFrequency, labels = sortedDoctorFrequency, pos = 3, cex = 0.7)


```


```{r}

# Loading the required libraries
library(tidyverse)
library(ggplot2)

# Loading the doctorwho dataset
doctorWhodf <- read.csv("doctorwho.csv", header = TRUE, sep = ",")

# Computing the frequency of each doctor
doctorFrequency <- table(doctorWhodf$FavoriteDoctor)

# Sorting the frequency in descending order
# sortedDoctorFrequency <- sort(doctorFrequency, decreasing = TRUE)

# Creating a vector of colors for each doctor
doctorColors <- c("blue", "red", "black", "orange")

# Creating the pie chart for the data
pie(doctorFrequency, 
    labels = names(doctorFrequency), 
    col = doctorColors, 
    main = "Patient's Visiting Frequency per Doctors",
    cex = 0.8, 
    font.main = 1, 
    font.lab = 1, 
    font.axis = 1)

# Adding the grid lines
# grid(lty = 3, lwd = 0.5, col = "gray")

# Adding the legend to the pie chart
legend("right", cex = 0.8, legend = names(doctorFrequency), fill = doctorColors)


```



## Q2. A Simple Time Series
The standard R base comes with a multivariate time series data set for major European stock indices from 1991 to 1998 in "ts" format (EuStockMarkets).  Display a paneled plot of the four indices represented (hint:  Don't use ggplot2--consider R's built-in plotting functions).

```{r}

# Loading the EuStockMarkets dataset
data(EuStockMarkets)

# Creating a 2x2 plot panel for the four indices
par(mfrow = c(2, 2))

# Creating a plot for each of the four indices with different colors and axis limits
plot(EuStockMarkets[, "DAX"], type = "l", main = "Germany (DAX)", col = "#999999", ylab = "Index Value", xlab = "Year", ylim = c(0, ceiling(max(EuStockMarkets[, "DAX"]))))
plot(EuStockMarkets[, "SMI"], type = "l", main = "Switzerland (SMI)", col = "#E69F00", ylab = "Index Value", xlab = "Year", ylim = c(0, ceiling(max(EuStockMarkets[, "SMI"]))))
plot(EuStockMarkets[, "CAC"], type = "l", main = "France (CAC)", col = "#56B4E9", ylab = "Index Value", xlab = "Year", ylim = c(0, ceiling(max(EuStockMarkets[, "CAC"]))))
plot(EuStockMarkets[, "FTSE"], type = "l", main = "UK (FTSE)", col = "#D55E00", ylab = "Index Value", xlab = "Year", ylim = c(0, ceiling(max(EuStockMarkets[, "FTSE"]))))


```


## Q3. Third item
A certain measurement tool collects statistics about server performance once a day and reports the instantaneous measure at that moment.  The following data set represents the measured number of requests per second for each day in 2014: (http://cs.ucf.edu/~wiegand/ids6938/datasets/RequestData.csv)
Plot a scatter-plot of the continuous time series data.  Fit a smoothed curve to the data over the year, as well. 
Hint:  You can convert the data string to a date type (which will help with plotting the x-axis) as follows:
```R
... aes(x=strptime(Day,"%m-%d-%Y") ...
```

```{r}
# Loading the required libraries
library(tidyverse)
library(ggplot2)

# Loading the RequestData dataset
requestDatadf <- read.csv("RequestData.csv", header = TRUE, sep = ",")
# head(requestDatadf)

# Converting the "Day" column to POSIXct object
requestDatadf$Day <- as.POSIXct(requestDatadf$Day, format="%m-%d-%Y")

# Creating the scatter-plot of the continuous time-series data
ggplot(requestDatadf, aes(x=Day, y=RequestsPerSecond)) + 
  geom_point(size = 1.5, alpha = 0.6, color = "blue") +
  labs(title = "Number of Requests per Second per Day in 2014",
       x = "Date", 
       y = "Requests per Second") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank())

# Creating the scatter-plot of the continuous time-series data and fitting a smoothed curve
ggplot(requestDatadf, aes(x=Day, y=RequestsPerSecond)) + 
  geom_point(color = "blue") +
  geom_smooth(method = "loess", se = TRUE, color = "red") +
  labs(title = "Number of Requests per Second per Day in 2014 with Curve Fitting",
       x = "Date", 
       y = "Requests per Second") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank())


```



## Q4. UCF Building Energy Usage Time Series
Data for 5 UCF building structures are in the data folder of this repository.

* Read in the datasets into R, and list what data is being captured for each building.


```{r}

# Loading the libraries
library(tidyverse)

# Loading the 5 UCF Building Structures datasets

# Reading the UCF Biological Sciences building dataset 
biologicalSciencesdf <- read.csv("2021-UCF Biological Sciences.csv", header = TRUE, sep = ";")

# Listing the data that is being captured for the Biological Sciences building
str(biologicalSciencesdf)
head(biologicalSciencesdf)
#print(biologicalSciencesdf)
summary(biologicalSciencesdf)

# Reading the UCF CFE Federal Credit Union Arena building dataset 
cfeCreditUnionArenadf <- read.csv("2021-UCF CFE Federal Credit Union Arena.csv", header = TRUE, sep = ";")

# Listing the data that is being captured for the CFE Federal Credit Union Arena building
str(cfeCreditUnionArenadf)
head(cfeCreditUnionArenadf)
#print(cfeCreditUnionArenadf)
summary(cfeCreditUnionArenadf)

# Reading the UCF Milican Hall building dataset 
milicanHalldf <- read.csv("2021-UCF Milican Hall.csv", header = TRUE, sep = ";")

# Listing the data that is being captured for the Milican Hall building
str(milicanHalldf)
head(milicanHalldf)
#print(milicanHalldf)
summary(milicanHalldf)

# Reading the UCF Student Union building dataset 
studentUniondf <- read.csv("2021-UCF Student Union.csv", header = TRUE, sep = ";")

# Listing the data that is being captured for the Student Union building
str(studentUniondf)
head(studentUniondf)
#print(studentUniondf)
summary(studentUniondf)

# Reading the UCF Trevor Colbourn Hall building dataset 
trevorColbournHalldf <- read.csv("2021-UCF Trevor Colbourn Hall.csv", header = TRUE, sep = ";")

# Listing the data that is being captured for the Trevor Colbourn Hall building
str(trevorColbournHalldf)
head(trevorColbournHalldf)
#print(trevorColbournHalldf)
summary(trevorColbournHalldf)


```


* Plot a grouped bar graph of the timeseries Electricity data for each building, with a 7-day sliding window line graph overlayed. Plot a stacked bar graph, with a 7-day sliding window, of all the energy usage of the buildings in one graph.



```{r}
# Loading the libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(zoo)
library(lubridate)

# Loading the 5 UCF Building Structures datasets

# Reading the UCF Biological Sciences building dataset 
biologicalSciencesdf <- read.csv("2021-UCF Biological Sciences.csv", header = TRUE, sep = ";")


# Reading the UCF CFE Federal Credit Union Arena building dataset 
cfeCreditUnionArenadf <- read.csv("2021-UCF CFE Federal Credit Union Arena.csv", header = TRUE, sep = ";")

# Reading the UCF Milican Hall building dataset 
milicanHalldf <- read.csv("2021-UCF Milican Hall.csv", header = TRUE, sep = ";")

# Reading the UCF Student Union building dataset 
studentUniondf <- read.csv("2021-UCF Student Union.csv", header = TRUE, sep = ";")

# Reading the UCF Trevor Colbourn Hall building dataset 
trevorColbournHalldf <- read.csv("2021-UCF Trevor Colbourn Hall.csv", header = TRUE, sep = ";")



```


```{r}
# Loading the libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(zoo)
library(lubridate)

# Loading the 5 UCF Building Structures datasets

# Reading the UCF Biological Sciences building dataset 
biologicalSciencesdf <- read.csv("2021-UCF Biological Sciences.csv", header = TRUE, sep = ";")

# Reading the UCF CFE Federal Credit Union Arena building dataset 
cfeCreditUnionArenadf <- read.csv("2021-UCF CFE Federal Credit Union Arena.csv", header = TRUE, sep = ";")

# Reading the UCF Milican Hall building dataset 
milicanHalldf <- read.csv("2021-UCF Milican Hall.csv", header = TRUE, sep = ";")

# Reading the UCF Student Union building dataset 
studentUniondf <- read.csv("2021-UCF Student Union.csv", header = TRUE, sep = ";")

# Reading the UCF Trevor Colbourn Hall building dataset 
trevorColbournHalldf <- read.csv("2021-UCF Trevor Colbourn Hall.csv", header = TRUE, sep = ";")


```



* Do a decomposition of each data field for the five buildings (hint: stl), plot both weekly and monthly time periods.

```{r}
# Loading the libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(zoo)
library(lubridate)

# Loading the 5 UCF Building Structures datasets

# Reading the UCF Biological Sciences building dataset 
biologicalSciencesdf <- read.csv("2021-UCF Biological Sciences.csv", header = TRUE, sep = ";")

head(biologicalSciencesdf)

# biologicalSciencesdf$Electric <- as.numeric(biologicalSciencesdf$Electric, na.rm = TRUE)

# Performing decomposition of each data field for Biological Sciences building using stl
# biologicalSciencesdfDecomposed <- stl(biologicalSciencesdf$Electric, s.window = "periodic")


```



```{r}


```


* Plot the buildings' electricity on a radar plot.


```{r}
# Loading the libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(zoo)
library(lubridate)
library(fmsb)

# Loading the 5 UCF Building Structures datasets

# Reading the UCF Biological Sciences building dataset 
biologicalSciencesdf <- read.csv("2021-UCF Biological Sciences.csv", header = TRUE, sep = ";")

# Reading the UCF CFE Federal Credit Union Arena building dataset 
cfeCreditUnionArenadf <- read.csv("2021-UCF CFE Federal Credit Union Arena.csv", header = TRUE, sep = ";")

# Reading the UCF Milican Hall building dataset 
milicanHalldf <- read.csv("2021-UCF Milican Hall.csv", header = TRUE, sep = ";")

# Reading the UCF Student Union building dataset 
studentUniondf <- read.csv("2021-UCF Student Union.csv", header = TRUE, sep = ";")

# Reading the UCF Trevor Colbourn Hall building dataset 
trevorColbournHalldf <- read.csv("2021-UCF Trevor Colbourn Hall.csv", header = TRUE, sep = ";")


# Creating a data frame with the electricity data for each building
datadf <- data.frame(
  Biological_Sciences = as.numeric(biologicalSciencesdf$Electric),
  CFE_Credit_Union_Arena = as.numeric(cfeCreditUnionArenadf$Electric),
  Milican_Hall = as.numeric(milicanHalldf$Electric),
  Student_Union = as.numeric(studentUniondf$Electric),
  Trevor_Colbourn_Hall = as.numeric(trevorColbournHalldf$Electric)
)


# data <- as.data.frame(datadf)
# dim(data)

# Creating a vector of labels for the radar plot
labels <- c("Biological Sciences", "CFE Credit Union Arena", "Milican Hall", "Student Union", "Trevor Colbourn Hall")




```



* Plot the buildings' electricity on 5 different heat maps, and 1 for the overall electric usage.


```{r}
# Load the ggplot2 package
library(ggplot2)

# Loading the 5 UCF Building Structures datasets

# Reading the UCF Biological Sciences building dataset 
biologicalSciencesdf <- read.csv("2021-UCF Biological Sciences.csv", header = TRUE, sep = ";")

# Reading the UCF CFE Federal Credit Union Arena building dataset 
cfeCreditUnionArenadf <- read.csv("2021-UCF CFE Federal Credit Union Arena.csv", header = TRUE, sep = ";")

# Reading the UCF Milican Hall building dataset 
milicanHalldf <- read.csv("2021-UCF Milican Hall.csv", header = TRUE, sep = ";")

# Reading the UCF Student Union building dataset 
studentUniondf <- read.csv("2021-UCF Student Union.csv", header = TRUE, sep = ";")

# Reading the UCF Trevor Colbourn Hall building dataset 
trevorColbournHalldf <- read.csv("2021-UCF Trevor Colbourn Hall.csv", header = TRUE, sep = ";")

# Check the column names of each data frame
names(biologicalSciencesdf)
names(cfeCreditUnionArenadf)
names(milicanHalldf)
names(studentUniondf)
names(trevorColbournHalldf)

# Check the structure of each data frame
str(biologicalSciencesdf)
str(cfeCreditUnionArenadf)
str(milicanHalldf)
str(studentUniondf)
str(trevorColbournHalldf)



```


